Aggressive Scheduling for Numerical Programs

### 2012-05-02 report/draft.tex
added experimental evaluation.

%% NOTE: need to explain this WAY better in the future.
For our experiments we benchmarked multiple variants of Matrix-Vector(GEMV) on
several systems with different microarchitectures. The variants included a
baseline in which no loops were unrolled, two unrolled variants where the
inner-most loop was fully unrolled and the other dimension was unrolled by
either 1 or 4. Additionally, there were two variants that were software pipelined
versions of the unrolled code. The two different unrollings resulted in
completely different schedules in the software pipelined code. It is worth
noting that we specifically chose our input sizes to fit entirely in the L1
cache, so that performance would mostly be impacted by the number of
instruction stalls.

@@

\begin{verbatim}
+ What algorithms are we using?
+ What systems are we using?

@@

\end{verbatim}

>>>>

%% NOTE: need to explain this WAY better in the future.
For our experiments we benchmarked multiple automatically generated variants
of Matrix-Vector(GEMV) on several systems with different
microarchitectures. The variants included a baseline in which no loops were
unrolled, two unrolled variants where the inner-most loop was fully unrolled
and the other dimension was unrolled by either 1 or 4. Additionally, there
were two variants that were software pipelined versions of the unrolled
code. The two different unrollings resulted in completely different schedules
in the software pipelined code. It is worth noting that we specifically chose
our input sizes to fit entirely in the L1 cache, so that performance would
mostly be impacted by the number of instruction stalls. Lastly, we were using
4-way 32bit floating point vector instructions.

@@

\begin{comment}
+ What algorithms are we using?
+ What systems are we using?

@@

As anticipated, the use of software pipelining provided a substantial
performance gain over unrolling alone, and on the more modern system (Core i7,
Fig.~\ref{fig:gemv_corei7}) yielded an impressive 1.8x speedup over unrolling
alone. The other two out-of-order systems (Fig.~\ref{fig:gemv_core2} and
Fig.~\ref{fig:gemv_xeon}) saw similar gains. The reason for performance
improvement over simple unrolling is attributed to the fact that software
pipelining breaks dependency chains and provides the processor's ready
instruction window with a variety of independent instructions to choose from,
thus reducing the risk of a dependency stall.

%% NOTE: need to explain the delay thing better
Interestingly, software pipelining did not provide as significant of a gain on
the Intel Atom (Fig.~\ref{fig:gemv_atom}) as it did on the out-of-order
systems. This could be attributed to the fact that we were using this
processor in 32bit mode and thus we were limited to half the number of
registers that are available in 64bit mode. Furthermore, our current software
pipelining implementation does not support long delays, which might have been
necessary on this system.
